---
title: "Kelp Project Test"
author: "Jessica French"
date: "2023-01-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(here)

#Set up data directory
data_dir <- here("..", "GIMMS3G_NDVI_TRENDS_1275", "data")
```


## First tutorial using `ncdf4` package following [this](https://rpubs.com/boyerag/297592) tutorial. 

```{r}
# load in packages

library(ncdf4)

```


first the output of the steps in the tutorial. 


```{r}
# read in data used in tutorial
tutorial_data <- nc_open(file.path(data_dir, 'gimms3g_ndvi_1982-2012.nc4'))

# print the metadata
print(tutorial_data)
```


Next the out put from the kelp file 
```{r}
# Read in the netCDF file downloaded form the lter 
kelp_data <- nc_open(file.path(data_dir, 'LandsatKelpBiomass_2022_Q4_withmetadata.nc'))

# print the metadata
print(kelp_data)
```
The rest of the outputs for the tutorial are shown on the webpage

```{r}
# From the metadata it looks like the location data is store in the station variable. Extract station variable and look at it 
 station <- ncvar_get(kelp_data, "station")

print(station)
```
My interpretation of the situation is that the location information is is linked to the station variable. Each observation only has information for the station and each lat and lon has an associated station. Need to join the lat lon information to the observations based on station. 


