---
title: "California EEZ Vis"
output: html_document
date: "2023-01-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goals: I want to make a visualization of California's EEZ and the United States EEZ out to 150 meters.

Need: us_states (sp_data or sp_data_large), bathymetry file(local)

Plan: add a 3 mile and 200 mile buffer to us_states, rasterize, clip to cells with 150 meters of depth or less.

```{r label= "Load Packages"}
# Load packages

library(tidyverse)
library(spData)
library(terra)
library(sp)
library(sf)
library(tmap)
library(here)

```


## Data Sources

Federal Maritime Boundaries Data is downloadable from [NOAA](https://nauticalcharts.noaa.gov/data/us-maritime-limits-and-boundaries.html)

California Coastal Zone is downloadable from [ca.gov](https://map.dfg.ca.gov/metadata/ds0990.html)

Polygons of US states is availabel through the `spData` package. 

Bathymetry data is available...

```{r label= "load in data"}

#Load in data

## Bathymetry
depth_rast <- rast(here("..", "depth.tif"))|> 
  project("EPSG:4326")

## US states polygons
states <- us_states |> 
  st_transform(crs = 4326)

## Maritime boundaries from NOAA
boundaries <- st_read(here("..", "maritime_boundaries", "USMaritimeLimitsNBoundaries.shp")) |> 
  st_transform(crs = 4326)

## Coastal Zone

coastal_zone <- st_read(here("..", "cali_coast_zone", "ds990.gdb"))|> 
  st_transform(crs = 4326)

# Sediment thickness

sediment_depth <- rast(here("..", "SedimentThickness_SantaBarbaraChannel", "SedimentThickness_SantaBarbaraChannel.tif")) |> 
  project("EPSG:4326")
  

### Check that CRS is the same for all datasets

crs(states) == crs(depth_rast)
crs(states) == crs(boundaries)
crs(states) == crs(coastal_zone)
crs(states) == crs(sediment_depth)


st_crs(boundaries, parmeters = TRUE)
crs(depth_rast)
```

```{r label= "filter states to California"}
#filter states to California
## This will be used in the next chunk to filter boundaries
cali <- states |> 
  filter(NAME == "California")

```

```{r label= "filter boundaries to pacific coast}
# Boundaries to Pacific Coast
pacific_boundaries <- boundaries |> 
  filter(REGION == "Pacific Coast") 
```

```{r }
depth_rast_reclass <- clamp(depth_rast, lower = -150, upper= 0, value=FALSE)
```

```{r label= "initial vis"}
tmap_mode(mode = "view")

map <- 
  tm_shape(sediment_depth) +
  tm_raster() +
  tm_shape(pacific_boundaries) +
  tm_lines() +
  tm_shape(coastal_zone) +
  tm_borders(col = "purple2", 
             lwd = 2) +
  # tm_shape(depth_rast_reclass) +
  # tm_raster()+
  tm_scale_bar()
map
  
```
