---
title: "kelp nitrate climate data"
output: html_document
date: "2023-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load in packages
library(tidyverse)
library(here)
library(sf)
library(tmap)
library(lubridate)
```


## LTER dataset downloaded from [here](https://sbclter.msi.ucsb.edu/data/catalog/package/?package=knb-lter-sbc.144). Nitrogen concentration is in micromoles per liter of seawater at the surface. Temporal scale is quarter and year from 1987 to 2019.

```{r}
# load in data on kelp nitrate and climate factors downloaded from lters website, filter to years 2014 and later

lter_data <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nitrogen_Data/Nitrate_LTER/kelp_no3_waves_quarterly_long (1).csv") |> 
  filter(year >= 2014)

# load in the location data

lter_locations <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nitrogen_Data/Nitrate_LTER/locations (1).csv")

```

```{r}
# Join location data to nitrogen data by site id. 
# Convert the lat and lon columns into point geometry. 
##### I don't know if wgs84 is the right crs for this area but it will work for now.
lter_data_geom <- left_join(lter_data, lter_locations, by = "site_id") |> 
  st_as_sf(coords = c(lon = "lon", lat = "lat"), crs = "EPSG:4326")
```

```{r}
# Visualize using tmap
tmap_mode(mode = "view")

tm_shape(lter_data_geom) +
  tm_dots(col = "no3")
```

## Cal COFI nitrogen dataset downloaded from [here](https://calcofi.org/data/oceanographic-data/bottle-database/). There are two files, cast and bottle. The cast file contains the metadata including the location and dates that the samples were taken. The bottle data is the actual measurements were are interested in but we need to join the cast data to it to have the location. 

```{r}
# Load in both the cast(metadata).. 

# Filter the cast data to exclude anything from before 2014. Select columns to have the lat, lon, cast count(Cst_Cnt), and quarter.
# Convert lat lon columns into point geometry
CalCOFI_cast <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nitrogen_Data/CalCOFI_Database_194903-202001_csv_22Sep2021/194903-202001_Cast.csv") |> 
  filter(Year >= 2014) |> 
  select(c("Cst_Cnt", "Quarter", "Year", "Lat_Dec", "Lon_Dec")) 
```

```{r}
# Load in bottle data and join the cast data to it. There is no date to filter by in the bottle data so the join acts as a filter.  I am not sure what the locale=.. is doing exactly but the file wouldn't read without it.
CalCOFI_bottle <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nitrogen_Data/CalCOFI_Database_194903-202001_csv_22Sep2021/194903-202001_Bottle.csv", locale=locale(encoding="latin1")) |> 
  right_join(CalCOFI_cast, by = "Cst_Cnt")|> 
  st_as_sf(coords = c(lon = "Lon_Dec", lat = "Lat_Dec"), crs = "EPSG: 4326")


```

There are parsing problems in columns 57, 59, 60, 61, 62. None are variables we are interested in. 
```{r label = "CalCOFI"}
# Column 57 is DIC2
CalCOFI_bottle[57]
# Column 59 is TA2
CalCOFI_bottle[59]
# Column 60 is pH1
CalCOFI_bottle[60]
# Column 61 is pH2
CalCOFI_bottle[61]
# Column 62 is DIC quality comment. 
CalCOFI_bottle[62]
```

```{r label = "CalCOFI viz"}
# Create map to visualize the area the data cover
tm_shape(CalCOFI_bottle) +
  tm_dots(col = "NO3uM")
```

## Plumes and Blooms, data available to download [here](http://www.oceancolor.ucsb.edu/plumes_and_blooms/) data exploration, this does not have the same spatial coverage but has more data in the area is does cover compared to the CalCOFI data. This data set is also extremely lacking in metadata so I need to reach out to the researchers to see if they have it.

```{r label = Plumes and Blooms}
# Read in the plumes and blooms csv
# Convert the date column to a data so we can filter by year, for now I am assuming YMD for the date format but need to get ahold of the researchers before committing to that. Convert lat and lon into point geom.
PlumesBlooms <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nitrogen_Data/Plumes_and_Blooms_bottle_file_data.csv", na = c('-999.00000', '-999', "")) |>
  mutate("DATE" = lubridate::ymd(DATE),
         "LON" = LON*-1,) |>
  filter(DATE >= 2014-01-01,
         LON < 100,
         LAT >= 30 & LAT <= 40) |>
  st_as_sf(coords = c(lon = "LON", lat = "LAT"), crs = "EPSG:4326")
```


There is definitely some wonkyness with the coordinates: NAs longitude in the latitude column, all the longitudes are on the wrong side of the planet, and values that do not make sense. Again need to reach out and get metadata from them before actually incorporating this in project. Also there should not be negative values for nitrogen concentration :( 

```{r label = "Plumes and Blooms Viz"}
# Visualize PlumesBlooms data
tmap_mode(mode = "view")
tm_shape(PlumesBlooms) +
  tm_dots(col = "NO3NO2")
```


## Some of the data comes form bio-oracle who have an r package that we will check out below. The NOAA mean nitrogen concentration data was derived from bio oracle data that we can access directly through their website and an r package that they maintain. They resolution isnt awesome but I think we could use this to improve coverage. 

```{r}
# Read in the library
library(sdmpredictors)

# List the datasets
datasets <- list_datasets(terrestrial = FALSE, marine = TRUE)
list_layers(datasets)

# There is a lot here and we need to look into what might be useful.
```

