---
title: "kelp nitrate climate data"
output: html_document
date: "2023-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load in packages
library(tidyverse)
library(here)
library(sf)
library(tmap)
```


## LTER dataset downloaded from [here](https://sbclter.msi.ucsb.edu/data/catalog/package/?package=knb-lter-sbc.144). Nitrogen concentration is in micromoles per liter of seawater at the surface. Temporal scale is quarter and year from 1987 to 2019.

```{r}
# load in data on kelp nitrate and climate factors downloaded from lters website

lter_data <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nitrogen_Data/Nitrate_LTER/kelp_no3_waves_quarterly_long (1).csv")

# load in the location data

lter_locations <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nitrogen_Data/Nitrate_LTER/locations (1).csv")

```

```{r}
# Join location data to nitrogen data by site id. 
# Convert the lat and lon columns into point geometry. 
##### I don't know if wgs84 is the right crs for this area but it will work for now.
lter_data_geom <- left_join(lter_data, lter_locations, by = "site_id") |> 
  st_as_sf(coords = c(lon = "lon", lat = "lat"), crs = "EPSG:4326")
```

```{r}
# Visualize using tmap
tmap_mode(mode = "view")

tm_shape(lter_data_geom) +
  tm_dots(col = "no3")
```

## Cal COFI nitrogen dataset downloaded from [here](https://calcofi.org/data/oceanographic-data/bottle-database/). There are two files, cast and bottle. The cast file contains the metadata including the location and dates that the samples were taken. The bottle data is the actual measurements were are interested in but we need to join the cast data to it to have the location. 

```{r}
# Load in both the cast(metadata).. 

# Filter the cast data to exclude anything from before 2014. Select columns to have the lat, lon, cast count(Cst_Cnt), and quarter.
# Convert lat lon columns into point geometry
CalCOFI_cast <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nitrogen_Data/CalCOFI_Database_194903-202001_csv_22Sep2021/194903-202001_Cast.csv") |> 
  filter(Year >= 2014) |> 
  select(c("Cst_Cnt", "Quarter", "Year", "Lat_Dec", "Lon_Dec")) 
```

```{r}
# Load in bottle data and join the cast data to it. There is no date to filter by in the bottle data so the join acts as a filter. 
CalCOFI_bottle <- read_csv("/Users/jfrench/Documents/MEDS/Capstone/DATA/Nitrogen_Data/CalCOFI_Database_194903-202001_csv_22Sep2021/194903-202001_Bottle.csv", locale=locale(encoding="latin1")) |> 
  right_join(CalCOFI_cast, by = "Cst_Cnt")|> 
  st_as_sf(coords = c(lon = "Lon_Dec", lat = "Lat_Dec"), crs = "EPSG: 4326")


```

There are parsing problems in columns 57, 59, 60, 61, 62. None are variables we are interested in. 
```{r}
# Column 57 is DIC2
CalCOFI_bottle[57]
# Column 59 is TA2
CalCOFI_bottle[59]
# Column 60 is pH1
CalCOFI_bottle[60]
# Column 61 is pH2
CalCOFI_bottle[61]
# Column 62 is DIC quality comment. 
CalCOFI_bottle[62]
```

```{r}
# Create map to visualize the area the data cover
tm_shape(CalCOFI_bottle) +
  tm_dots(col = "NO3uM")
```

